On the adhesivity of EGGS

### 2025-02-13 calco/main.tex
pass biblio

In this section we briegly recall the notion \emph{$\X$-hypergraph}. In order to do it is necessary to have a monad $(-)^\star$ playing a role analogous to the usual \emph{Kleene star}, also known as \emph{list monad}, $(-)^\star\colon \Set\to \Set$ sending a set to the free monoid on it \cite{sakarovitch2009elements,wadler1995monads}. We will recall some of its proprieties.

@@

Let $\mathcal{G}_F$ be the resulting hypergraph. One can now show that  sending $F$ to $\mathcal{G}_F$ can be extended to an equivalence $\mathcal{G}_{-}\colon \Set^{\catname{H}}\to \hyp$ (see \cite{castelnovo2023thesis, castelnovo2022newcriterionmathcalmmathcalnadhesivity} for details).

@@

This characterization, in turn, provides us with the following result \cite{castelnovo2022newcriterionmathcalmmathcalnadhesivity,castelnovo2023thesis}.

@@

\Cref{thm:slice-functors}, \Cref{lem:reg} and \Cref{prop:push} allow us to recover the following result, previously proved by direct computation in \cite[Thm.~4.2]{corradini2005term}.
\begin{corollary}
	The category $\tg$ is quasiadhesive.
\end{corollary}

@@

\begin{lemma}\label[lemma]{colim}
	Let $L\colon \A\to \X$ and $R\colon \B\to \X$ be functors and $F\colon \D\to\comma{L}{R}$ be a diagram such that $L$ preserves colimits along $U_L\circ F$. Then the family $\{U_L, U_R\}$ jointly creates colimits of $F$ (see \cite{castelnovo2023thesis,castelnovo2022newcriterionmathcalmmathcalnadhesivity}).
\end{lemma}
\begin{proof}
	Suppose that $U_L\circ F$ and $U_R\circ F$ have colimiting cocones  $\left(A, \{a_D\}_{D\in \D}\right)$ and $\left(B, \{b_D\}_{D\in \D}\right)$ respectively. By hypothesis $\left(L(A), \{L\left(a_D\right)\}_{D\in \D}\right)$ is colimiting for $L\circ U_L\circ F$. Now, if we define
	\[F(D):=(A_D, B_D, f_D)\]
	then we have arrows $R(a_i)\circ f_D\colon L(A_D)\to R(B)$ that forms a cocone on $L\circ U_L\circ F$: if $d\colon D\to D'$ is an arrow in $\D$ then $F(d)$ is an arrow in $\comma{L}{R}$ and so
	\begin{align*}
		R\left(b_{D'}\right)\circ f_{D'}\circ L(U_L(F(d)))&=R\left(b_{D'}\right)\circ R\left(U_R\left(F(d)\right)\right)\circ f_D\\&=R\left(b_{D'}\circ U_R\left(F(d)\right)\right)\circ f_D\\&=R\left(b_D\right)\circ f_D
	\end{align*}
	Thus there exists $f\colon L(A)\rightarrow R(B)$ such that
	\[\xymatrix@C=35pt{ L(A_D) \ar[r]^{L(a_D)} \ar[d]_{f_D}& L(A) \ar[d]^{f}\\ R(B_D) \ar[r]_{R(b_D)} &R(B)}\]
	Notice that $f$ is the unique arrow in $\X$ wich makes $\left(a_D, b_D\right)$ an arrow $\left(A_D, B_D, f_D\right)\to \left(A, B, f\right)$ of $\comma{L}{R}$. If we show that $\left((A, B, f), \left\{(a_D, b_D)\right\}_{D\in \D}\right)$ is colimiting for $F$ we are done.

@@

\section{Comparison with Ghica}
\todo{Lo vogliamo fare?}

>>>>

In this section we briefly recall the notion \emph{$\X$-hypergraph}. In order to do it is necessary to have a monad $(-)^\star$ playing a role analogous to the usual \emph{Kleene star}, also known as \emph{list monad}, $(-)^\star\colon \Set\to \Set$ sending a set to the free monoid on it \cite{sakarovitch2009elements,Wadler95}. We will recall some of its proprieties.

@@

Let $\mathcal{G}_F$ be the resulting hypergraph. One can now show that  sending $F$ to $\mathcal{G}_F$ can be extended to an equivalence $\mathcal{G}_{-}\colon \Set^{\catname{H}}\to \hyp$ (see \cite{castelnovo2023thesis,CastelnovoGM24} for details).

@@

This characterization, in turn, provides us with the following result \cite{CastelnovoGM24,castelnovo2023thesis}.

@@

\Cref{thm:slice-functors}, \Cref{lem:reg} and \Cref{prop:push} allow us to recover the following result, previously proved by direct computation in \cite[Thm.~4.2]{CorradiniG05}.
\begin{corollary}
	The category $\tg$ is quasiadhesive.
\end{corollary}

@@

\begin{lemma}\label[lemma]{colim}
	Let $L\colon \A\to \X$ and $R\colon \B\to \X$ be functors and $F\colon \D\to\comma{L}{R}$ be a diagram such that $L$ preserves colimits along $U_L\circ F$. Then the family $\{U_L, U_R\}$ jointly creates colimits of $F$ (see \cite{castelnovo2023thesis,CastelnovoGM24}).
\end{lemma}
\begin{proof}
	Suppose that $U_L\circ F$ and $U_R\circ F$ have colimiting cocones  $\left(A, \{a_D\}_{D\in \D}\right)$ and $\left(B, \{b_D\}_{D\in \D}\right)$ respectively. By hypothesis $\left(L(A), \{L\left(a_D\right)\}_{D\in \D}\right)$ is colimiting for $L\circ U_L\circ F$. Now, if we define
	\[F(D):=(A_D, B_D, f_D)\]
	then we have arrows $R(a_i)\circ f_D\colon L(A_D)\to R(B)$ that forms a cocone on $L\circ U_L\circ F$: if $d\colon D\to D'$ is an arrow in $\D$ then $F(d)$ is an arrow in $\comma{L}{R}$ and so
	\begin{align*}
		R\left(b_{D'}\right)\circ f_{D'}\circ L(U_L(F(d)))&=R\left(b_{D'}\right)\circ R\left(U_R\left(F(d)\right)\right)\circ f_D\\&=R\left(b_{D'}\circ U_R\left(F(d)\right)\right)\circ f_D\\&=R\left(b_D\right)\circ f_D
	\end{align*}
	Thus there exists $f\colon L(A)\rightarrow R(B)$ such that
	\[\xymatrix@C=35pt{ L(A_D) \ar[r]^{L(a_D)} \ar[d]_{f_D}& L(A) \ar[d]^{f}\\ R(B_D) \ar[r]_{R(b_D)} &R(B)}\]
	Notice that $f$ is the unique arrow in $\X$ wich makes $\left(a_D, b_D\right)$ an arrow $\left(A_D, B_D, f_D\right)\to \left(A, B, f\right)$ of $\comma{L}{R}$. If we show that $\left((A, B, f), \left\{(a_D, b_D)\right\}_{D\in \D}\right)$ is colimiting for $F$ we are done.

@@

%\section{Comparison with Ghica}
%\todo{Lo vogliamo fare?}
